---
title: "Statistical Inference Peer Assessment Part 2 of 2"
author: "Joe Cannon"
date: "October 19, 2015"
output: html_document
---

```{r  echo=FALSE}
suppressWarnings(suppressMessages(library(knitr)))
```

## Effect of different Viamin C delivery methods in guinea pig tooth growth

### Overview
Dataset:
The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

This part of the project I will
- Load the ToothGrowth data and perform some basic exploratory data analyses 
- Provide a basic summary of the data.
- Used confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
- State my conclusions and the assumptions needed for your conclusions. 


#### Load libraries
```{r echo=FALSE}
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(datasets)))
suppressWarnings(suppressMessages(library(gridExtra)))
suppressWarnings(suppressMessages(library(grid)))
suppressWarnings(suppressMessages(library(gtable)))

```

####1.) Load ToothGrowth data
```{r LoadData} 
data(ToothGrowth)
toothGrowth <- ToothGrowth 
toothGrowth$dose <- as.numeric(toothGrowth$dose) # convert to factor
toothGrowth$len <- as.numeric(toothGrowth$len)
```


####2.) Basic Summary of the data
```{r, fig.Height= 6, fig.Width =10}
str(toothGrowth)
summary(toothGrowth)
head(toothGrowth)
```

Create and average length for each dose, and suppliment
```{r average ,fig.width=10, fig.height=5}
ATG <- aggregate(len~dose+supp,toothGrowth,mean)

  mytheme <- gridExtra::ttheme_default(
      core = list(fg_params=list(cex = 1.3)),
      colhead = list(fg_params=list(cex = 1.5)),
      rowhead = list(fg_params=list(cex = .5)))
  
  table <- gridExtra::tableGrob(ATG, theme = mytheme,rows = NULL)
  title <- textGrob("Supliment Dose/Tooth Growth",gp=gpar(fontsize=10,fontface="bold"))

  footnote <- textGrob(date() , x=0, hjust=0,
                       gp=gpar( fontface="italic",fontsize=5),just = 0)
  padding <- unit(.5,"line")
  table <- gtable_add_rows(table, 
                           heights = grobHeight(title) + padding,
                           pos = 0)
  table <- gtable_add_rows(table, 
                           heights = grobHeight(footnote)+padding)

  
  table <- gtable_add_grob(table, list(title,footnote),
                           t=c(1, nrow(table)), l=c(1,2), 
                           r=ncol(table))


  grid.newpage()
  grid.draw(table)
```

This is the result of the average tooth growth based on dose of supp.

```{r BasicSummary}
ggplot(ATG, aes(x=dose,y=len,colour=supp),size=2) +
    geom_line() +
    ggtitle("Tooth length by dose of Supliment")+
    xlab("Dose")+
    ylab("Average Tooth Length")      
```

###3.) Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r CalcConfInt}
test1 <- t.test(len~supp, paired=F, var.equal=T, data=ToothGrowth[which(ToothGrowth$dose==.5),])
test2 <- t.test(len~supp, paired=F, var.equal=F, data=ToothGrowth[which(ToothGrowth$dose==.5),])


result <- data.frame("conf"=c(attr(test1$conf,"conf.level"),attr(test2$conf,"conf.level")),
                     "p_value"=c(test1$p.value, test2$p.value),
                     "conf_low"=c(test1$conf[1],test2$conf[1]),
                     "conf_high"=c(test1$conf[2],test2$conf[2]),
                      row.names=c("Equal Var","Unequal Var"))
plow <- min(result$p_value)
phigh <- max(result$p_value) 

```


```{r plot_ttest results}
  mytheme <- gridExtra::ttheme_default(
      core = list(fg_params=list(cex = .8)),
      colhead = list(fg_params=list(cex = .8)),
      rowhead = list(fg_params=list(cex = .8)))
  
  table <- gridExtra::tableGrob(result, theme = mytheme)
  title <- textGrob("T Confidence Tests",gp=gpar(fontsize=20))

  footnote <- textGrob("Paired does not make sense in this test as the guinea pigs are supposedly different. ", x=0, hjust=0,
                       gp=gpar( fontface="italic",fontsize=8))
  padding <- unit(1,"line")
  table <- gtable_add_rows(table, 
                           heights = grobHeight(title) + padding,
                           pos = 0)
  table <- gtable_add_rows(table, 
                           heights = grobHeight(footnote)+padding)
  
  table <- gtable_add_grob(table, list(title,footnote),
                           t=c(1, nrow(table)), l=c(1,2), 
                           r=ncol(table))

  grid.newpage()
  grid.draw(table)
```



####4.) State your conclusions and the assumptions needed for your conclusions.

###Conclusion
Based on the data provided, it can be stated that both VC, and OJ have beneficial results with regards to tooth length in guinea pigs with a p-Value of `r plow` to `r phigh`. However at smaller doses Orange Juice seems to have a greater impact than  ascorbic acid. The difference in the delivery methods diminishes to 0 as doses increase. This is a blended test of all doses. 

Assumptions:
*There is a coorilation of  tooth length in guinea pigs to vitamin C 
*This is a random sample of guinea pigs
*60 is a large enough sample size 



